### 2024-05-26 getting-started.qmd
20240525 - rushing EPA by age; ypc by age

load(file = here("data", "nfl_players.RData"))
load(file = here("data", "nfl_teams.RData"))
load(file = here("data", "nfl_rosters.RData"))
load(file = here("data", "nfl_rosters_weekly.RData"))
load(file = here("data", "nfl_schedules.RData"))
load(file = here("data", "nfl_combine.RData"))
load(file = here("data", "nfl_draftPicks.RData"))
load(file = here("data", "nfl_depthCharts.RData"))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = here("data", "nfl_actualStats_weekly.RData"))
load(file = here("data", "nfl_injuries.RData"))
load(file = here("data", "nfl_snapCounts.RData"))
load(file = here("data", "nfl_espnQBR.RData"))
load(file = here("data", "nfl_nextGenStats_weekly.RData"))
load(file = here("data", "nfl_advancedStatsPFR.RData"))
load(file = here("data", "nfl_playerContracts.RData"))
load(file = here("data", "nfl_ftnCharting.RData"))
load(file = here("data", "nfl_playerIDs.RData"))
load(file = here("data", "nfl_rankings.RData"))
load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))
```

@@

# Merge with player weekly stats
player_age_stats_offense <- left_join(
  player_age_offense,
  nfl_actualStats_game_offense_weekly,
  by = c(c("season","week","player_id","game_id")))

player_age_stats_defense <- left_join(
  player_age_defense,
  nfl_actualStats_game_defense_weekly,
  by = c(c("season","week","player_id","game_id")))

player_age_stats_kicking <- left_join(
  player_age_kicking,
  nfl_actualStats_game_kicking_weekly,
  by = c(c("season","week","player_id","game_id")))

player_age_stats_offense$years_of_experience <- as.integer(player_age_stats_offense$years_of_experience)
player_age_stats_defense$years_of_experience <- as.integer(player_age_stats_defense$years_of_experience)
player_age_stats_kicking$years_of_experience <- as.integer(player_age_stats_kicking$years_of_experience)
```

@@

>>>>

# Downloaded Data
load(file = here("data", "nfl_players.RData"))
load(file = here("data", "nfl_teams.RData"))
load(file = here("data", "nfl_rosters.RData"))
load(file = here("data", "nfl_rosters_weekly.RData"))
load(file = here("data", "nfl_schedules.RData"))
load(file = here("data", "nfl_combine.RData"))
load(file = here("data", "nfl_draftPicks.RData"))
load(file = here("data", "nfl_depthCharts.RData"))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = here("data", "nfl_actualStats_weekly.RData"))
load(file = here("data", "nfl_injuries.RData"))
load(file = here("data", "nfl_snapCounts.RData"))
load(file = here("data", "nfl_espnQBR.RData"))
load(file = here("data", "nfl_nextGenStats_weekly.RData"))
load(file = here("data", "nfl_advancedStatsPFR.RData"))
load(file = here("data", "nfl_playerContracts.RData"))
load(file = here("data", "nfl_ftnCharting.RData"))
load(file = here("data", "nfl_playerIDs.RData"))
load(file = here("data", "nfl_rankings.RData"))
load(file = here("data", "nfl_expectedFantasyPoints_weekly.RData"))
load(file = here("data", "nfl_expectedFantasyPoints_offense.RData"))

# Calculated Data
load(file = here("data", "nfl_actualStats_career.RData"))
load(file = here("data", "nfl_actualStats_seasonal.RData"))
```

@@

# Merge with player weekly stats
player_age_stats_offense <- left_join(
  player_age_offense %>% select(-position, -position_group),
  nfl_actualStats_game_offense_weekly,
  by = c(c("season","week","player_id","game_id")))

player_age_stats_defense <- left_join(
  player_age_defense %>% select(-position, -position_group),
  nfl_actualStats_game_defense_weekly,
  by = c(c("season","week","player_id","game_id")))

player_age_stats_kicking <- left_join(
  player_age_kicking %>% select(-position, -position_group),
  nfl_actualStats_game_kicking_weekly,
  by = c(c("season","week","player_id","game_id")))

player_age_stats_offense$years_of_experience <- as.integer(player_age_stats_offense$years_of_experience)
player_age_stats_defense$years_of_experience <- as.integer(player_age_stats_defense$years_of_experience)
player_age_stats_kicking$years_of_experience <- as.integer(player_age_stats_kicking$years_of_experience)

# Merge player info with seasonal stats
player_seasonal_offense <- left_join(
  nfl_actualStats_offense_seasonal,
  nfl_players %>% select(-position, -position_group, -season),
  by = c("player_id" = "gsis_id")
)

player_seasonal_defense <- left_join(
  nfl_actualStats_defense_seasonal,
  nfl_players %>% select(-position, -position_group, -season),
  by = c("player_id" = "gsis_id")
)

player_seasonal_kicking <- left_join(
  nfl_actualStats_kicking_seasonal,
  nfl_players %>% select(-position, -position_group, -season),
  by = c("player_id" = "gsis_id")
)

# Calculate age
season_startdate <- nfl_schedules %>% 
  group_by(season) %>% 
  summarise(startdate = min(gameday, na.rm = TRUE))

player_seasonal_offense <- player_seasonal_offense %>% 
  left_join(
    season_startdate,
    by = "season"
  )

player_seasonal_defense <- player_seasonal_defense %>% 
  left_join(
    season_startdate,
    by = "season"
  )

player_seasonal_kicking <- player_seasonal_kicking %>% 
  left_join(
    season_startdate,
    by = "season"
  )

player_seasonal_offense$age <- interval(
  start = player_seasonal_offense$birth_date,
  end = player_seasonal_offense$startdate
) %>% 
  time_length(unit = "years")

player_seasonal_defense$age <- interval(
  start = player_seasonal_defense$birth_date,
  end = player_seasonal_defense$startdate
) %>% 
  time_length(unit = "years")

player_seasonal_kicking$age <- interval(
  start = player_seasonal_kicking$birth_date,
  end = player_seasonal_kicking$startdate
) %>% 
  time_length(unit = "years")
```

@@

### Rushing Yards per Carry By Player Age {#sec-plotRushEPAbyAge}

```{r}
# Prepare Data
rushing_attempts <- nfl_pbp %>% 
  dplyr::filter(
    season_type == "REG") %>% 
    filter(
      rush == 1,
      rush_attempt == 1,
      qb_scramble == 0,
      qb_dropback == 0,
      !is.na(rushing_yards))

rb_yardsPerCarry <- rushing_attempts %>% 
  group_by(rusher_id, season) %>% 
  summarise(
    ypc = mean(rushing_yards, na.rm = TRUE),
    rush_attempts = n(),
    .groups = "drop") %>% 
  ungroup() %>% 
  left_join(
    nfl_players %>% select(-season),
    by = c("rusher_id" = "gsis_id")
  ) %>% 
  filter(
    position_group == "RB",
    rush_attempts >= 50) %>% 
  left_join(
    season_startdate,
    by = "season"
  )

rb_yardsPerCarry$age <- interval(
  start = rb_yardsPerCarry$birth_date,
  end = rb_yardsPerCarry$startdate
) %>% 
  time_length(unit = "years")

# Create Plot
ggplot2::ggplot(
  data = rb_yardsPerCarry,
  ggplot2::aes(
    x = age,
    y = ypc)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth() +
  ggplot2::labs(
    x = "Rushing Back Age (years)",
    y = "Rushing Yards per Carry/season",
    title = "2023 NFL Rushing Yards Per Carry per Season by Player Age",
    subtitle = "(minimum 50 rushing attempts)"
  ) +
  ggplot2::theme_classic()
```

```{r}
# Subset Data
rb_seasonal <- player_seasonal_offense %>% 
  filter(position_group == "RB")

# Create Plot
ggplot2::ggplot(
  data = rb_seasonal,
  ggplot2::aes(
    x = age,
    y = rushing_epa)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth() +
  ggplot2::labs(
    x = "Rushing Back Age (years)",
    y = "Rushing EPA/season",
    title = "2023 NFL Rushing EPA per Season by Player Age"
  ) +
  ggplot2::theme_classic()
```
